\chapter{Sets and Measure Theory	}

\begin{multicols}{2}[\subsubsection*{Contents of this chapter}]
   \printcontents{}{1}{\setcounter{tocdepth}{2}}
\end{multicols}



Sets are a term to describe collections of things. The things could be countable objects, such as the integers between $1$ and $10$, or contain a continuum, for example all the real numbers between $1$ and $10$. Unless otherwise specified, the elements of a set are assumed to be distinct and as not having an internal order. That is, the set of letters in "Mississippi" is $\{M,i,s,p \} = \{i,p,M,s 	\} = ... $ and so on. A concise resource for notation is \citeasnoun{stanfordsetnotation}. 


% representation
\section{Representation}

\begin{tabular}{ll}
Statement Form & $\{\mathrm{integers\ between\ 1\ and\ 5}\}$\\
Roster Form & $\{1,2,3,4,5\}$ \\
Setbuilder Form & $\{x|x\in\mathbb{N}, 1\leq x \leq 5 \} $
\end{tabular}


% Set Properties, Types of Sets
\section{Set Properties and Types of Sets}

\subsection{Cardinality}
The cardinality of a set is a measure of the size of a set. For finite sets, the cardinality is simply the number of elements. For example, the set  $A = \{a,b,c,d\}$ has size $|A|=4$.

For infinite sets, this intuition breaks down, though it is still possible to make meaningful comparisons. 

\subsection{$\emptyset$ Empy Set, Null Set}
The empty set is the set with no elements. It is denoted $\{\}$ or $\emptyset$. 

\subsection{Singleton Set}
A singleton set is a set with one element. I.e., $\{a\}, \{b\}$ are singleton subsets of $\{a,b\}$. A singleton has cardinality 1.

\subsection{Infinite Sets}

\subsection{Multisets}
A multiset is a set that may contain an element more than once. I.e. it makes sense to write $A=\{a,a,a,b,b\}$. The number of times an element is included is the \textit{multiplicity}. The multiset $A$ may be defined in terms of a \textit{multiplicity function} $m_A(x)$, which describes the multiplicity of a type of element $x\in U$ where $U$ may be referred to as the \textit{universe} (en lieu of saying "univseral"!). The multiplicity function allows the extension of set characteristics and operations to multisets.

\subsubsection{Support}
The support of a multiset is:

\begin{equation}
Supp(A) = \{ x\in U | 	m_A(x)>0\}
\end{equation} 

Which is the set of distinct elements in $A$. I.e., if $A = \{a,a,a,b,b,c\}$ then $Supp(A) = \{a,b,c\}$.

\subsubsection{Cardinality}

The cardinality is given by:

\begin{equation}
|A| = \sum_{x\in U} m_A(x)
\end{equation}

\subsubsection{$\subseteq$ Inclusion}
The concept of subsets can be extended to multisets as:

\begin{equation}
A \subseteq B
\end{equation}

if 

\begin{equation}
\forall x\in U, \ \ m_A(x) \leq m_B(x)
\end{equation}


\subsubsection{$\bigcap$ Intersection}
The intersection of multisets is sometimes called the \textit{infimum} or \textit{greatest common divisor}. If $A\cap B = C$, then $C$ has multiplicity function:

\begin{equation}
m_C(x) = min\left(m_A(x),m_B(x)\right)
\end{equation}

That is, it is like an elementwise minimum.

\subsubsection{$\bigcup$ Union}
In the context of multisets, the term \textit{union} sometimes refers to multiset addition. Otherwise, it should refer to the overlap of two multisets, i.e.:
 
\begin{equation}
m_C(x) = max\left(m_A(x),m_B(x)\right)
\end{equation}

That is, it is like an elementwise maximum.

\subsubsection{$\uplus$ Multiset Addition} 
In contrast to sets in which distinct elements are only contained once, multiset addition makes sense and uses a special symbol "$\uplus$". If $A\uplus B = C$, then $C$ has multiplicity function:

\begin{equation}
m_C(x) = m_A(x) + m_B(x)
\end{equation} 

\subsubsection{Multiset Subtraction}
Multisets can be subtracted, with the condition that the multiplicity can not be less than zero. If $A- B = C$, then $C$ has multiplicity function:

\begin{equation}
m_C(x) = max\left(m_A(x) - m_B(x), 0\right)
\end{equation}

% Powerset
\subsection{Powersets}
The powerset $P(S)$ of a set $S$ is the set of all subsets of $S$, including the empty set and the set itself. That is:

\begin{equation}
P(S) = \{ A : A\subseteq S\}
\end{equation}

My guess is it's called powerset because the number of subsets of S, $|P(S)| = 2^{|S|}$. I have also seen the notation $2^S$ to refer to the space of subsets of $S$.



% Measurable Sets
\subsection{Measurable Sets}
Measurable sets have a way of measuring volume on them in a non-trivial way. That is, for some subset of a measurable set $A_i \subseteq A$, it is possible to define a metric $\mu$ so that $\mu(A_i) \neq0)$. Measurable sets are the elements of a $\sigma$-algebra. 

% universal set
\subsection{Universal Set}
The \textit{universal set} is understood to refer to something that is not allowed in the context of Russell's Paradox. It is nevertheless useful to define a set $S$ so that $S^c = \emptyset$ and I've seen this type of set referred to as \textit{universal set} a couple of times. In the context of probability theory, $S$ may be the whole event space.




%% open sets, closed sets
\subsection{Open Sets, Closed Sets}
Open sets and closed sets are generalizations of open and closed intervals on the real line. Open sets a mentioned plenty in the context of measure theory. A set is open if and only if its complement is closed, and a set is closed if and only if its complement is open. Counterintuitively, a set can be both open and closed at the same time, or it can be neither open nor closed at the same time.

\subsubsection{Interior Points}
An interior point of a set $A$ is any point $X$ so that for some $\epsilon > 0$, the open interval $X-\epsilon,X+\epsilon \subseteq A$ is contained within $A$. One might imagine an interior point as a point that has a neighborhood (a "ball") that is included in $A$. 

\subsubsection{Accumulation Points} 
In contrast, accumulation points do not have a neighborhood contained in $A$. Formally, for some $\epsilon>0$, $(X-\epsilon,X+\epsilon)\cap (A\setminus \{X\}) \neq \emptyset$, which is only possible for a point that is on the very boundary of $A$.

\subsubsection{Open Sets}
Unsurprisingly, an open set is a set that only has interior points, i.e. $\{X: X\in A, \exists \epsilon > 0 \mathrm{\ s.\ th.\ } X-\epsilon,X+\epsilon \subseteq A  \}$.

\subsubsection{Closed Sets}
Closed sets contain all of their accumulation points. That is, the boundary is included. Points might also be standing alone, for example:

\begin{equation}
[2,4]\cup\{1\}
\end{equation}

Is a closed set.

\subsubsection{Both Open and Closed, Neither Open Nor Closed}
The set of real numbers $\mathbb{R}$ is both open and closed. Any point on the real line has a neighborhood in $\mathbb{R}$, so it is open. On the other hand, the accumulation points of $\mathbb{R}$ are $\mathbb{R}$, so it is closed. An interval that has some but not all accumulation points is neither open nor closed. For example $[1,2)$.	

%% Image
\subsection{Image}
\label{sec:image}
The image are the elements of the codomain that a given subset of the domain is mapped to. For example, given a function $f:\mathbb{R} \rightarrow \mathbb{R}: f(x) = x^2$, the image of $\{-3,-2,2,3\}$ in the domain are the elements $\{4,9\}$ of the codomain.

%% Preimage 
\subsection{Preimage, Inverse Image}
\label{sec:preimage}
The preimage are elements of the domain that are mapped to a given subset of the codomain. For example, given a function $f:\mathbb{R} \rightarrow \mathbb{R}: f(x) = x^2$, the preimage of the elements $\{4,9\}$ of the codomain are the elements $\{-3,-2,2,3\}$ of the domain..


%% Convex Sets
\subsection{Convex Sets}
Convex sets are subsets of vector spaces in which any point along the line connecting to points within the set is contained within the set. A disk in $\mathbb{R}^2$ is a convex set. A crescent in $\mathbb{R}^2$ is not a convex set. The boundary of a convex set is a convex function. For example, a parabola $f(x) = x^2$ with $x\in\mathbb{R}$ is a convex function, and the area above the parabola, called the epigraph, is a convex subset of $\mathbb{R}$. Convex optimization deals with the optimization of convex functions over convex sets. The intersection of convex sets is always convex, but the union of convex sets is only convex under certain conditions. 

More generally, if $S$ is a convex set, then affine combinations of the elements $\mathbf{x}_i  \in S$ of the form:

\begin{equation}
\sum_i \mathbf{x}_i \lambda_i 
\label{eq:convexsets}
\end{equation}  

With $\sum_i \lambda_i = 1$ and $\lambda_i\geq 0$ are also contained in $S$. Rather than just the points along the line between two points, these are the essentially the weighted averages of multiple points in the set.  

\subsubsection{Example: Discrete Probability Distributions}
The set of discrete probability distributions $P = \{ \mathbf{p}=(p_1,p_2,...): ||\mathbf{p}||_1 = 1, p_i \geq 0\}$ is a convex set. Property \ref{eq:convexsets} means that a weighted average over members of $P$ is also a member of $P$, so long as the weights satisfy $\sum_i \lambda_i = 1$. In general, this guarantees that marginalization results in a probability. I.e. if $p(X|Y) = \mathbf{p}(X|Y) \in P$ and $p(Y) = \mathbf{p}(Y) \in P$, then $\mathbf{p}(X) = \sum_i p_i(X|Y) p_i(Y) \in P$. 

%% Choice sets, Transverse Sets, Cross-sections
\subsection{Choice Sets, Transversal Sets, Cross-Sections}
\label{sec:choicesets}
Choice sets, transversal sets or cross-sections are sets that are assembled by picking exactly one element from each member of a family of disjoint sets. The axiom of choice (cf. section \ref{sec:axiomofchoice}) states that such a set can always be formed.

\subsubsection{Example: Integers}
Consider the partition of the integers between $1$ and $40$ into a family of disjoint sets that each contain $10$ elements: 

\begin{equation}
S = \{ \{1,...,10\}, \{11,...,20\}, \{21,...,30\},\{31,...,40\} \}
\end{equation}

Then a transversal set $A$ could for example be formed by picking the smallest number in each of the subsets. 

\begin{equation}
A = \{1,11,21,31\}
\end{equation}

Though we might have picked any other choice function.  


% Set operations
\section{Set Operations}

\subsection{$A^c$ Complement}
Given a subset $A \subseteq X$, the complement of the subset $A^c$ refers to everything that is not contained in $A$, i.e. $A^c = X\setminus A$. 


\subsection{$\bigcup$ Union}
The union of sets is the set that contains all of their elements, counting all elements only once. The cardinality of the union is calculated using the important inclusion exclusion principle (\ref{sec:inclusionexclusion}). In terms of logic, think $A\cup B$ is "A or B (or both)".

\subsection{$\bigcap$ Intersection}
The intersection of sets is the elements shared by all sets. $A\cap B$ is "A and B".

% Disjoint Union
\subsection{$\bigsqcup$ Disjoint Union, Discriminated Union}
The disjoint union, or discriminated union, of sets is the union formed in a way in which the information about which subset an element belonged to is preserved. One way to write this is to include a subset index with each element, i.e.:

\begin{equation}
\bigsqcup_{i\in\{i\}}A_i = \bigcup_{i\in\{i\}} \{(x,i): x\in A_i\}
\end{equation}

For example, for the two sets $A=\{1,2,3,4\}$ and $B=\{1,2,3\}$, the disjoint union is:

\begin{equation}
A\sqcup B = \{(1,A),(2,A),(3,A),(4,A),(1,B),(2,B),(3,B)\}
\end{equation}

The cardinality of the disjoint union is simply the sum:

\begin{equation}
\left|\bigsqcup_{i\in\{i\}}A_i\right| = \sum_{i\in\{i\}}|A_i
\end{equation}

I have also seen the disjoint union be used in the context of forming the union of disjoint sets, perhaps to stress that the sets are disjoint and the cardinality can be calculated through simple summation as above. For example, in dividing up the interval: $[0,2]=[0,1)\sqcup[1,2]$.


% Bijection Principle
\section{Bijection Principle}
The bijection principle states that when a bijection exists between the elements of two sets, then those sets have the same size. This principle is very useful in combinatorics, and it allows for meaningful comparisons of infinite sets.


% DeMorgan
\section{DeMorgan's Rules}
\label{sec:demorgan}
DeMorgan's Rules relate the complement of the union to the intersection of the complements, and the complement of the intersection to the union of the complements.

\begin{equation}
\left(\bigcup_{i\in\{i\}}A_i\right)^c = \bigcap_{i\in\{i\}}A^c_i
\end{equation}

\begin{equation}
\left(\bigcap_{i\in\{i\}}A_i\right)^c = \bigcup_{i\in\{i\}}A^c_i
\end{equation}


%Inclusion Exclusion
\section{Inclusion - Exclusion Principle}
\label{sec:inclusionexclusion}

The inclusion-exclusion principle is used to calculate the size of the union of sets. This requires counting each region of some complicated overlapping Venn diagram exactly once, which, in turn requires accounting for overcounting wherever sets overlap. Let $\{A_i | i\in \{i\}_n \}$ be a collection of $n$ overlapping sets indexed by $i\in \{i\}_n$, then the inclusion-exclusion principle is given by:

\begin{equation}
\left|\bigcup_{i\in\{i\}_n} A_i\right| = \sum^n_{k=1} (-1)^{k-1} \sum_{\{j\}_k \subseteq \{i\}_n} \left|\bigcap_{j\in\{j\}_k} A_j\right|
\end{equation}


Where the sum over $\{j\}_k \subseteq \{i\}_n$ is over all $k$-element subsets of $\{i\}_n$.

\subsection{Example: n=2 Sets and n=3 Sets}

\subparagraph{n=2}
\begin{equation}
\begin{array}{rl}
\left|\bigcup_{i\in\{1,2\}}A_i\right| =&  \sum^2_{k=1} (-1)^{k-1} \sum_{\{j\}_k \subseteq \{1,2\}} \left|\bigcap_{j\in\{j\}_k} A_j\right|\\
=&(-1)^{0}\left(|A_1| + |A_2|\right) \\
&+ (-1)^{1}\left(|A_1 \cap A_2| \right)
\end{array}
\end{equation}


\subparagraph{n=3}
\begin{equation}
\begin{array}{rl}
\left|\bigcup_{i\in\{1,2,3\}}A_i\right| =&  \sum^3_{k=1} (-1)^{k-1} \sum_{\{j\}_k \subseteq \{1,2,3\}} \left|\bigcap_{j\in\{j\}_k} A_j\right|\\
=&(-1)^{0}\left(|A_1| + |A_2| + |A_3|\right) \\
&+ (-1)^{1}\left(|A_1 \cap A_2|  + |A_1 \cap A_3| + |A_2 \cap A_3|  \right) \\ 
&+ (-1)^{2}\left(|A_1 \cap A_2 \cap A_3|   \right)
\end{array}
\end{equation}

\subsubsection{Example: Counting Integers}

How many integers are there between 1 and 100 that are neither divisible by 3,5 nor 7?

Let $S$ be the set of all integers between 1 and 100. The size of the set is $|S| = 100$. The subset of $S$ that is numbers divisible by 3 is $A_3 \subseteq S$ with $|A_3| = 33$ because $100/3 = 33.\overline{333}$. Similarly, $|A_5| = 20$ and $|A_7| = 14$.  The set of integers that is not divisible by 3, 5 or 7 is:

\begin{equation}
S \setminus \bigcup_{i\in\{3,5,7\}} A_i
\end{equation}

So that the sought after quantity is :

\begin{equation}
\begin{array}{rl}
\left|S \setminus \bigcup_{i\in\{3,5,7\}} A_i\right| =& |S| - \left[ |A_3| + |A_5| + |A_7| \right.\\
& \left. - |A_3 \cap A_5| - |A_3 \cap A_7| - |A_5\cap A_7| + |A_3\cap A_5\cap A_7|\right]
\end{array}
\end{equation}

The size of the intersection $|A_3\cap A_5| = 6$ because 100 is 6 times divisible by $3\times 5 = 15$. Similarly, $|A_3 \cap A_7| =  4$, $|A_5 \cap A_7| =  2$ and $|A_3\cap A_5 \cap A_7| =  0|$. Hence:

\begin{equation}
\left|S \setminus \bigcup_{i\in\{3,5,7\}} A_i\right| = 100 - 33 - 20 - 14 + 6 + 4 + 2 - 0 = 45
\end{equation}

There are 45 integers between 1 and 100 that are not divisible by 3, 5 or 7.


\section{Axiom of Choice}
\label{sec:axiomofchoice}
The axiom of choice simply states that, given a collection of nonempty, mutually disjoint sets, it is possible to assemble a \textit{transversal} or \textit{choice} set that consists of exactly one element from each of the sets in the collection. For example, consider the students in 1st, 2nd, 3rd.. etc grades at a school to be a collection of nonempty, mutually disjoint sets. Then the axiom of choice says that it is possible to assemble a subset of students with exactly one student from each grade. 


In terms of functions, one might think of defining a \textit{choice function} on the family of mutually disjoint sets, which selects members from the collection of sets and adds them to the \textit{choice} set. According to the axiom of choice, a choice function can be defined for any collection of nonempty, mutually disjoint subsets. This implies that all surjective functions have a right inverse. That is, for any surjective function $f:X\rightarrow Y$, there exists a function $g:Y\rightarrow X$ so that $f(g(y)) = y$.   

The axiom of choice turns out to be associated with famous names and deep consequences \cite{stanfordaxiomofchoice}.

\subsection{Example: Pairs of Real Numbers, Right Inverse}
The collection of rank-2 sets of pairs of real numbers $A= \{ A_x = \{x,-x\} : x\in \mathbb{R}^{+}\}$ are a collection of mutually disjoint subsets. A transversal set $B$ may be assembled by choosing the largest element of the tuple: $B = \{y : y max(A_x), A_x \in A \}$. An equivalent surjection $f:A\rightarrow B$ is $f(A_x) = x$. There is a right-inverse $g:B\rightarrow A$ which is $g(x) = (x,-x)$ so that $f(g(x)) = x$.







\input{./chapters/sections/sets_sigma.tex}